Capon Beamforming Applied on Moving Objects - An Analysis of Local Lateral Shift-Invariance

### 2013-11-30 HRUIofMovingObjects/jrnl_article.tex
Moving objects: Done some writing. Updated and added figures.

\begin{abstract}
%\boldmath
If an ultrasound imaging system provides a presentation of a moving object which is sensitive to local spatial shifts, the system is said to be local shift-invariant. This can for instance happen if the axial or lateral sampling is insufficient. The Capon beamformer has been shown to provide increased lateral resolution in ultrasound images. Increased lateral resolution should demand denser lateral sampling. In previous literature on Capon beamforming for medical ultrasound imaging however only single frame scenarios have been simulated, and temporal behaviour and effects caused by the increased resolution and lack of oversampling have therefore been neglected.

In this paper we investigate the local lateral shift-invariance of the Capon beamformer when imaging moving objects. We then discuss and investigate different methods in order to improve on it. We show that insufficient lateral sampling makes an imaging system based on the Capon beamformer locally lateral shift-variant. We then show that the locally lateral shift-invariance can be improved by oversampling based on phase rotation without affecting the acquisition frame rate and with minor changes in processing complexity.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

@@

In \cite{Cox1973}, Cox derives formulas for the resolution of the Capon beamformer, and presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. Cox states the following important fact about the Capon beamformer (the $\vec{k}_3$ processor): 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the steered response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to three times the system resolution ($1/32$ radians). Impinging on the array is a monochromatic plane wave originating from a source moving parallel to the array surface. The plot shows the steered response of the source when it is located at plus and minus half the system resolution. No spatially white noise present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) has also been added to the Capon beamformer output. A diagonal loading of $1/100$ corresponds to an signal-to-noise ratio of 20 dB before array gain is applied. The figure shows that a DAS beamformer, with a beamspacing equal to the system resolution, will yield a 3.92 dB scalloping loss \cite{Harris1978} if the source is moving parallel to the array. However if the Capon beamformer is used with the same beam density the scalloping loss could be as large as $31$ dB.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img fig1.eps}
}
\caption{Comparison of drop in output power in between receive beams. The steered response of both delay-and-sum (DAS) and Capon beamforming ($M=64$, $L=32$, $d=0.01$) is plotted for two different time instances. A monochromatic plane wave is located at an angle equal to half the system resolution, and is moved a distance corresponding to the system resolution for the second steered response.}
\label{fig:das_capon_beams}
\end{figure}

@@

In this paper we investigate the issue further, and we show that when a moving point scatterer is imaged with the Capon beamformer with the same beam spacing sufficient for DAS the predicted scalloping loss is observed, manifested as blinking point scatteres. The question is then, what can we do in order to improve on the situation without affecting the imaging frame rate and processing throughout. In this paper we therefore study the scalloping loss of the Capon beamforming and investigate different methods aiming at reducing this effect, and how the different methods will affect the imaging frame rate and the processing complexity.

In the next section we give an introduction to Capon beamforming for medical ultrasound imaging together with a short introduction to lateral sampling and shift invariance. In Section \ref{sec:capon_LLSI} the local lateral shift-invariance of the Capon beamformer is investigated, before the different methods are presented in Section \ref{sec:methods}. Then, in Section \ref{sec:res} and Section \ref{sec:dis} we present and discuss the results of applying the most promising method on \textit{in-vitro} data. Finally we give our conclusions in Section \ref{sec:con}.

@@

The lateral sampling required in an ultrasound image is given by the combined aperture size on transmit ($D_{tx}$) and receive ($D_{rx}$), and the center frequency ($f_c = c/\lambda$), where $c$ is the speed of sound in the body and $\lambda_c$ is the wavelength of the center frequency. The required Nyquist beam spacing is then \cite{Hergum2007}:
\begin{align}
\Delta < \frac{\lambda_c}{D_{tx} + D_{rx}}. \label{eq:resolution}
\end{align}
If this requirement is not fulfilled the scalloping loss could become visible in the image and the system will no longer be locally lateral shift-invariant. In an ultrasound image with 256 gray levels and 50 dB dynamic range, a 1 dB loss corresponds to 5 gray levels. This is approximately equal to the visibility threshold (Weber fraction of 2\%) for a grayscale image, and a loss larger than 1 dB could therefore end up being visible to the observer. In our analysis a system would, based on this reasoning, be regarded as locally lateral shift-invariant if the loss in between beams is smaller than 1 dB. However, depending on viewing light conditions and e.g. the selected gray level mapping, a larger loss might be acceptable.

In order to visualize the local lateral shift-invariance of a given imaging system, we will make use of lateral shift-variance plots (LSV-plots), introduced by Hergum \textit{et al.} \cite{Hergum2007}. A LSV-plot is constructed for a given imaging system by imaging a point scatterer moving laterally at constant speed. For each lateral position of the point scatterer the root-mean-square (RMS) of the image data is calculated in range, and the resulting beam profiles are stacked in to an image. For a given scan sequence there are twice as many point positions than there are receive lines. The LSV-plots in this paper are displayed as a contour plot with contours on -1, -2, -3, -6, -12 and -24 dB. The maximum for each beam profile is also marked with a black dot, and the amplitude variation among these points is presented in a subplot next to the LSV-plot.

@@

An imaging system is said to be locally lateral shift-invariant if the contours are diagonal and if the amplitude deviation among the maximum points is less than 1 dB. Then there is minimal geometric distortion (low P-MAE) and the scalloping loss will not be visible for the observer (low A-MAE). %Since both effects can occur at the same time it is important to have a measure not effected by geometric distortion.

Figure \ref{fig:das_shiftinvariant} presents a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). The red dotted vertical lines indicates the position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} with a probe setup equal to the M5S probe by GE Vingmed Ultrasound, Horten, Norway. The probe was excited with a 2.5 MHz and 1.5 cycles long pulse, and the lateral sampling density has been doubled prior to detection through coherent interpolation. We see that the contours are approximately diagonal and therefore we conclude that the system is locally lateral shift-invariant. The P-MAE and A-MAE are also minimal. In this, and all the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located exactly on this beam at one time instance.

The scalloping loss in Fig.\,\ref{fig:das_shiftinvariant} is smaller than what was predicted by Fig.\,\ref{fig:das_capon_beams}. To understand why note that Fig.\,\ref{fig:das_capon_beams} shows the one-way response of a passive array, and the system resolution has therefore been calculated based on the receive aperture only. Whereas the beamspacing in (\ref{eq:resolution}) is derived from the fact that the two-way lateral response of an active rectangular aperture is a $D_{tx}+D_{rx}$ wide triangle in K-space \cite{Hergum2009}. Even though the lateral bandwidth is increased on receive, the distance to the first zero (The Rayleigh criteria) remains the same when multiplying two sinc functions. Sampling according to (\ref{eq:resolution}) is therefore more than what is strictly needed in practise. The loss in Fig.\,\ref{fig:das_shiftinvariant} is because of this smaller than what was predicted.

In Fig.\,\ref{fig:das_shiftvariant} the beam density is reduced by a factor of two. We see how the scalloping loss is increased to a level where the contours no longer are diagonal. Hence, the imaging system is no longer locally lateral shift-invariant. We observe that the maximum peak is reduced when the point is located between two beams and the distance between the beams are too high. The loss in Fig.\,\ref{fig:das_shiftvariant} is smaller than the predicted loss in Fig\,\ref{fig:das_capon_beams} since there are only two samples per beam along the y-axis. Cubic interpolation of this result reveals a scalloping loss of 2.71 dB.

@@

\section{Local Lateral Shift-Invariance of the Capon Beamformer}\label{sec:capon_LLSI}
We will now investigate the local lateral shift-invariance of the Capon beamformer with parameters equal to the ones used in earlier work \cite{Synnevag2009, Asen}. In Fig.\,\ref{fig:capon1} the same point scatterer response as presented in Fig.\,\ref{fig:das_shiftinvariant} has been weighted with Capon weights calculated using (\ref{eq:capon_weights}) with $L = M/2 = 32$, $K=1$ and $d=1/100$. The set of lateral displacements is constructed in such a way that the center beam is the only one hitting the point scatterer exactly. We see that the scalloping loss is tremendous, and not far from the value predicted by Fig\,\ref{fig:das_capon_beams}. The difference between the predicted and the simulated loss might be caused by several factors. The most important one being that Fig.\,\ref{fig:das_capon_beams} shows a single frequency whereas the RMS operator used in Fig.\,\ref{fig:capon1} average a broadband signal in range. For a slice through the center of the pulse the scalloping loss is below -30dB.

The diagonal factor $d$ and the effective adaptive aperture size $L$ can be used to control the mainlobe width of the Capon beamformer. In both extreme cases, $d\gg1$ and $L=1$, the Capon beamformer becomes either a triangle or uniform-weighted DAS beamformer respectively, and will require a beamspacing according to (\ref{eq:resolution}). In Fig.\,\ref{fig:capon2} we show how the scalloping loss is reduced when the amount of diagonal loading is increased. We observe that Fig.\,\ref{fig:capon2} now contains a scaled version of the pattern presented in Fig.\,\ref{fig:das_shiftvariant}. From Fig.\,\ref{fig:capon} it should be obvious that the Capon beamformer either requires increased lateral sampling in order to make the imaging system locally lateral shift-invariant, or the resolution has to be reduced to the same level as DAS. It would be unwise to do all the computation involved with Capon beamforming with a high $d$ value and end up with a resolution equal to DAS. However, we see that the diagonal factor can be a useful tool if resolution has to be traded for the need of oversampling. The same is true for the subarray length parameter.

@@

\section{Methods}\label{sec:methods}
We will now discuss different ways of improving the local lateral shift-invariance when the Capon beamformer is used. From the previous section it should be clear that we somehow need to increase the lateral sampling to make the Capon imaging system locally lateral shift-invariant. In this section we will investigate how this oversampling should be conducted, how large it needs to be and how the proposed methods will affect the system frame rate and the processing complexity.

\subsection{Oversampling on Transmit}
The straight forward way of obtaining sufficient sampling in an active system is by increasing the number of transmit beams (tx-beams) until the system is laterally local shift-invariant. In Fig.\,\ref{fig:capon_oversampling_a} and Fig.\,\ref{fig:capon_oversampling_b} we have increased the lateral sampling on transmit by a factor of four and ten respectively. As expected the scalloping loss decreases with increased lateral sampling. Note that Fig.\,\ref{fig:capon_oversampling_b} is zoomed compared to Fig.\,\ref{fig:capon} in order to better see the beam-to-beam variation.

@@

Even though increasing the number of transmits improves the lateral shift-invariance it will also reduce the imaging frame rate with a factor equal to the oversampling factor. The processing complexity will also increase equally. For these two reasons, oversampling on transmit is something which is desirable to avoid in real time ultrasound imaging, and therefore we will not discuss this method further.

@@

\subsection{Oversampling by Parallel Receive Beamforming}
Parallel receive beamforming (PRB) means that multiple narrow receive beams (rx-beams) are constructed from one broad tx-beam. Then, by reducing the number of tx-beams and relying on multiple rx-beams in order to get sufficient sampling on receive, it is possible to achieve high frame rates in e.g. cardiac ultrasound imaging. However, PRB also introduces motion artifacts in the image which to some degree can be compensated for \cite{Hergum2007, Denarie2011, Denarie2013}. In \cite{Rabinovich2013} it is shown that synthetic transmit beamforming (STB) \cite{Hergum2007} works with Capon beamforming. In \cite{Asen} we also used STB in order to increase the beam density of a cardiac ultrasound recording. In the following discussion we consider PRB as being equal to oversampling on transmit when it comes to image quality, and the LSV-plots for PRB is therefore not included.

In cardiac ultrasound imaging, PRB is typically used in order to keep the number of rx-beams high while the number of tx-beams are reduced. When using PRB for oversampling we would like to maintain the same number of tx-beams while the number of rx-beams are increased. The acquisition frame rate will then not be affected. On the other hand more rx-beams will increase the number of covariance matrices, and thereby the number of matrix-inversions with a factor equal to the number of parallel receive lines. The processing times reported in \cite{Asen} will then also increase with the same factor. Since almost 10x oversampling is required in order to achieve lateral shift invariant imaging of a moving point scatterer, none of the implementations described in \cite{Asen} would be real time any more.

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since the ultrasound data are pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole sector by only using phase rotation. This is also possible in broadband applications as long as the phase rotation is limited to a fraction of the wave length (coarse-fine beamforming \cite{Thomenius}). It is therefore possible to make a sweep between the current beam and its two nearest neighbours using only phase rotation. Phase rotation is actually what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance edges.

@@

>>>>

\begin{abstract}
%\boldmath
If an ultrasound imaging system provides a presentation of a moving object which is sensitive to local spatial shifts, the system is said to be spatially locally shift-invariant. This can for instance happen if the axial or lateral sampling is insufficient. The Capon beamformer has been shown to provide increased lateral resolution in ultrasound images. Increased lateral resolution should demand denser lateral sampling. In previous literature on Capon beamforming for medical ultrasound imaging however only single frame scenarios have been simulated, and temporal behaviour and effects caused by the increased resolution and lack of oversampling have therefore been neglected.

In this paper we investigate the local lateral shift-invariance of the Capon beamformer when imaging moving objects. We then discuss and investigate different methods in order to improve on it. We show that insufficient lateral sampling makes an imaging system based on the Capon beamformer locally laterally shift-variant. We then show that the local lateral shift-invariance can be improved by oversampling based on phase rotation without affecting the acquisition frame rate and with minor changes in processing complexity.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

@@

In \cite{Cox1973}, Cox derives formulas for the resolution of the Capon beamformer, and presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. Cox states the following important fact about the Capon beamformer (the $\vec{k}_3$ processor): 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to two times the two-way system resolution ($1/64$ radians) using a narrow band pulse. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the scanned output when the object is located at plus and minus half the system resolution. No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) has also been added to the Capon beamformer output. A diagonal loading of $1/100$ corresponds to an signal-to-noise ratio of 20 dB before array gain is applied. The figure shows that a DAS beamformer, with a beamspacing equal to the system resolution, could yield a 1.9 dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However if the Capon beamformer is used with the same beam density the scalloping loss could be as large as $27$ dB.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img fig1_2way.eps}
}
\caption{Theoretical drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming ($M=64$, $L=32$, $d=0.01$) applied in a farfield-narrowband setting. The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located in the farfield of the scanned sector at an angle equal to half the system resolution, and for the latter the object is moved a distance corresponding to the system resolution.}
\label{fig:das_capon_beams}
\end{figure}

@@

In this paper we investigate the issue further, and we show that when a moving point scatterer is imaged with the Capon beamformer with the same beam spacing sufficient for DAS the predicted scalloping loss is observed, manifested as blinking point scatteres. The question is then, what can we do in order to improve on the situation without affecting the imaging frame rate and processing throughout. In this paper we therefore study the scalloping loss of the Capon beamforming and investigate different methods aimed at reducing it. We also discuss how the different methods will affect the imaging frame rate and the processing complexity.

In the next section we give an introduction to Capon beamforming for medical ultrasound imaging together with a short introduction to lateral sampling and spatial shift invariance. In Section \ref{sec:capon_LLSI} the local lateral shift-invariance of the Capon beamformer is investigated, before the different methods are presented in Section \ref{sec:methods}. Then, in Section \ref{sec:res} and Section \ref{sec:dis} we present and discuss the results of applying the most promising method on \textit{in-vitro} data. Finally we give our conclusions in Section \ref{sec:con}.

@@

The lateral sampling required in an ultrasound image is given by the combined aperture size on transmit ($D_{tx}$) and receive ($D_{rx}$), and the center frequency ($f_c = c/\lambda$), where $c$ is the speed of sound in the body and $\lambda_c$ is the wavelength of the center frequency. The required Nyquist beam spacing is then \cite{Hergum2007}:
\begin{align}
\Delta < \frac{\lambda_c}{D_{tx} + D_{rx}}. \label{eq:resolution}
\end{align}
If this requirement is not fulfilled the scalloping loss could become visible in the image and the system will no longer be locally laterally shift-invariant. In an ultrasound image with 256 gray levels and 50 dB dynamic range, a 1 dB loss corresponds to 5 gray levels. This is approximately equal to the visibility threshold (Weber fraction of 2\%) for a grayscale image, and a loss larger than 1 dB could therefore end up being visible to the observer. In our analysis a system would, based on this reasoning, be regarded as locally laterally shift-invariant if the loss in between beams is smaller than 1 dB. However, depending on viewing light conditions and e.g. the selected gray level mapping, a larger loss might be acceptable.

In order to visualize the local lateral shift-invariance of a given imaging system, we will make use of lateral shift-variance plots (LSV-plots), introduced by Hergum \textit{et al.} \cite{Hergum2007}. A LSV-plot is constructed for a given imaging system by imaging a point scatterer moving laterally at constant speed. For each lateral position the root-mean-square (RMS) of the image data is calculated in range, and the resulting beam profiles are stacked in to an image. For a given scan sequence there are twice as many point positions than there are receive lines. The LSV-plots in this paper are displayed as a contour plot with contours on -1, -2, -3, -6, -12 and -24 dB. The maximum for each beam profile is also marked with a black dot, and the amplitude variation among these points is presented in a subplot next to the LSV-plot.

@@

A given imaging system will then be locally laterally shift-invariant if the contours are diagonal and if the amplitude deviation among the maximum points is less than 1 dB. Then there is minimal geometric distortion (low P-MAE) and the scalloping loss will hopefully not be visible for the observer (low A-MAE). %Since both effects can occur at the same time it is important to have a measure not effected by geometric distortion.

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). The red dotted vertical lines indicates the position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} with a probe setup equal to the M5S probe by GE Vingmed Ultrasound, Horten, Norway. The probe was excited with a 2.5 MHz and 1.5 cycles long pulse with a $\text{f-number}=3$ and corresponding cosine apodization on receive. The transmit focus in azimuth and elevation was configured to be equal and the point scatterer was moved through this focus. The lateral sampling density has been doubled prior to detection through coherent interpolation. We see that the contours are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. In this, and all the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located exactly on this beam at one time instance.

The scalloping loss in Fig.\,\ref{fig:das_shiftinvariant} is smaller than what was predicted by Fig.\,\ref{fig:das_capon_beams}. To understand why note that Fig.\,\ref{fig:das_capon_beams} shows the scanned response from a narrowband system while the imaging system in Fig.\,\ref{fig:das_shiftinvariant} uses a broadband pulse. The beam in Fig.\,\ref{fig:das_shiftinvariant} while therefore be more similar to a squared gaussian function than the squared $sinc$ functions present in Fig.\,\ref{fig:das_capon_beams}. The imaging system in Fig.\,\ref{fig:das_shiftinvariant} also uses an expanding cosine window on receive.

%The scalloping loss in Fig.\,\ref{fig:das_shiftinvariant} is smaller than what was predicted by Fig.\,\ref{fig:das_capon_beams}. To understand why note that Fig.\,\ref{fig:das_capon_beams} shows the one-way response of a passive array, and the system resolution has therefore been calculated based on the receive aperture only. Whereas the beamspacing in (\ref{eq:resolution}) is derived from the fact that the two-way lateral response of an active rectangular aperture is a $D_{tx}+D_{rx}$ wide triangle in K-space \cite{Hergum2009}. Even though the lateral bandwidth is increased on receive, the distance to the first zero (The Rayleigh criteria) remains the same when multiplying two sinc functions. Sampling according to (\ref{eq:resolution}) is therefore more than what is strictly needed in practise. The loss in Fig.\,\ref{fig:das_shiftinvariant} is because of this smaller than what was predicted.

In Fig.\,\ref{fig:das_shiftvariant} the beam density is reduced by a factor of two. We see how the geometric distortion and scalloping loss is increased to a level where the contours no longer are diagonal and constant. Hence, the imaging system is no longer locally laterally shift-invariant. We observe that the maximum peak is both geometrically distorted and attenuated when the point is located between two beams and the distance between the beams are too high.

@@

\section{Local Lateral Shift-Invariance of the Capon Beamformer}\label{sec:capon_LLSI}
We will now investigate the local lateral shift-invariance of the Capon beamformer with parameters equal to the ones used in earlier work \cite{Synnevag2009, Asen}. In Fig.\,\ref{fig:capon1} the same point scatterer response as presented in Fig.\,\ref{fig:das_shiftinvariant} has been weighted with Capon weights calculated using (\ref{eq:capon_weights}) with $L = M/2 = 32$, $K=1$ and $d=1/100$. As mentioned, the set of lateral displacements is constructed in such a way that the center beam is the only one hitting the point scatterer exactly. The signal cancellation caused by the difference between the signal propagation vector and the assumed steering vector is therefore minimal at this location. From the plot we observe that the signal cancellation is high in all the other lateral positions. As a consequence the scalloping loss is tremendous, and not far from the predicted value in Fig\,\ref{fig:das_capon_beams}. %The difference between the predicted and the simulated loss might be caused by several factors. The most important one being that Fig.\,\ref{fig:das_capon_beams} shows a single frequency whereas the RMS operator used in Fig.\,\ref{fig:capon1} average a broadband signal in range. For a slice through the center of the pulse the scalloping loss is below -30dB.

The diagonal factor $d$ and the effective adaptive aperture size $L$ can be used to control the mainlobe width of the Capon beamformer. In both extreme cases, $d\gg1$ and $L=1$, the Capon beamformer becomes either a triangle or uniform-weighted DAS beamformer respectively, and will require a beamspacing according to (\ref{eq:resolution}). In Fig.\,\ref{fig:capon2} we show how the scalloping loss is reduced when the amount of diagonal loading is increased.

We observe that Fig.\,\ref{fig:capon2} now contains a scaled version of the pattern presented in Fig.\,\ref{fig:das_shiftvariant}. From Fig.\,\ref{fig:capon} it should be obvious that the Capon beamformer either requires increased lateral sampling in order to make the imaging system locally laterally shift-invariant, or the resolution has to be reduced to the same level as DAS. It would be unwise to do all the computation involved with Capon beamforming with a high $d$ value and end up with a resolution equal to DAS. However, we see that the diagonal factor can be a useful tool if resolution has to be traded for the need of oversampling. The same is true for the subarray length parameter.

@@

\section{Oversampling Methods}\label{sec:methods}
We will now discuss different ways of improving the local lateral shift-invariance when the Capon beamformer is used. From the previous section it should be clear that we somehow need to increase the lateral sampling to make the Capon imaging system locally laterally shift-invariant. In this section we will investigate how this oversampling should be conducted, how large it needs to be and how the proposed methods will affect the system frame rate and the processing complexity.

\subsection{Oversampling on Transmit}
The straight forward way of obtaining sufficient sampling in an active system is by increasing the number of transmit beams (tx-beams) until the system is locally laterally shift-invariant. In Fig.\,\ref{fig:capon_oversampling_a} and Fig.\,\ref{fig:capon_oversampling_b} we have increased the lateral sampling on transmit by a factor of four and ten respectively. As expected the geometric distortion and scalloping loss decreases with increased lateral sampling. Note that Fig.\,\ref{fig:capon_oversampling_b} is zoomed compared to Fig.\,\ref{fig:capon} in order to better see the beam-to-beam variation.

@@

Even though increasing the number of transmits improves the lateral shift-invariance it will also reduce the imaging frame rate with a factor equal to the oversampling factor. The processing complexity will also increase equally. For these two reasons, oversampling on transmit is something which is desirable to avoid in real time ultrasound applications, and therefore we will not discuss this method further.

@@

\subsection{Oversampling by Parallel Receive Beamforming}
Parallel receive beamforming (PRB) means that multiple narrow receive beams (rx-beams) are constructed from one broad tx-beam. Then, by reducing the number of tx-beams and relying on multiple rx-beams in order to get sufficient sampling on receive, it is possible to achieve high frame rates in e.g. cardiac ultrasound imaging. However, PRB also introduces motion artifacts in the image which to some degree can be compensated for \cite{Hergum2007, Denarie2011, Denarie2013}. In \cite{Rabinovich2013} it is shown that synthetic transmit beamforming (STB) \cite{Hergum2007} works with Capon beamforming. In \cite{Asen} we also used STB in order to increase the beam density of a cardiac ultrasound recording. In the following discussion we consider corrected PRB as being equal to oversampling on transmit when it comes to image quality. The LSV-plots for PRB are therefore not included.

In cardiac ultrasound imaging, PRB is typically used in order to keep the number of rx-beams high while the number of tx-beams are reduced. When using PRB for oversampling we would like to maintain the same number of tx-beams while the number of rx-beams are increased. The acquisition frame rate will then not be affected. On the other hand more rx-beams will increase the number of covariance matrices, and thereby the number of matrix-inversions with a factor equal to the number of parallel receive lines. The processing times reported in \cite{Asen} will then also increase with the same factor. Since more than ten times oversampling is required in order to achieve lateral shift invariant imaging of a moving point scatterer, none of the implementations described in \cite{Asen} would be real time any more.

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since the ultrasound data are pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole sector by only using phase rotation. This is also possible in broadband applications as long as the phase rotation is limited to a fraction of the wave length (coarse-fine beamforming \cite{Thomenius}). It is therefore possible to make a sweep between a given beam and its two nearest neighbours using only phase rotation without introducing to large errors. Phase rotation is actually what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance edges.

@@

The geometric distortion remaining in Fig. 7d should be possible to fix by a phase rotation version of STB.
